---
title: "Assignment 2 - Group 4"
author: "Katrin Grunert, Carmen Byrne Salsas, Bono Lardinois"
date: "14 October 2022"
output:
  html_document:
    df_print: paged
fontsize: 11pt
highlight: tango
---

`r sat = read.table("sat.txt",header=TRUE)`
`r tree = read.table("treeVolume.txt",header=TRUE)`

# Exercise 2.1

## Exercise 2.1.a

Text here

```{r,echo=FALSE,fig.height=3.5}
#code here
```

## Exercise 2.1.b

Text here

```{r, echo=FALSE}
#code here
```

## Exercise 2.1.c

Text here

```{r}
#code here
```
## Exercise 2.1.d

Text here

```{r}
#code here
```

# Exercise 2.2
## Exercise 2.2.a

Text here
```{r}
#code here
```

## Exercise 2.2.b

Text here

```{r}
#code here
```

## Exercise 2.2.c

Text here

```{r}
#code here
```


# Exercise 2.3
## Exercise 2.3.a

Text here

```{r}
#code here
```


## Exercise 2.3.b

Text here

```{r}
#code here
```


## Exercise 2.3.c

Text here

```{r}
#code here
```

# Exercise 2.4
## Exercise 2.4.a

Text here

```{r}
#code here
```

## Exercise 2.4.b

Text here

```{r}
#code here
```

# Exercise 2.5
## Exercise 2.5.a

Text here

```{r}
#code here
```

## Exercise 2.5.b

Text here

```{r}
#code here
```
