---
title: "Assignment 2 - Group 4"
author: "Katrin Grunert (13914103), Carmen Byrne Salsas (12968900), Bono Lardinois (11713364)"
date: "14 October 2022"
output:
  pdf_document: default
fontsize: 11pt
highlight: tango
---

```{r,echo=FALSE,fig.height=3.5}
sat = read.table("sat.txt",header=TRUE)
tree = read.table("treeVolume.txt",header=TRUE)
```

# Exercise 2.1

## Exercise 2.1.a

Text here

```{r,echo=FALSE,fig.height=3.5}
#code here
```

## Exercise 2.1.b

Text here

```{r, echo=FALSE}
#code here
```

## Exercise 2.1.c

Text here

```{r}
#code here
```
## Exercise 2.1.d

Text here

```{r}
#code here
```

# Exercise 2.2
## Exercise 2.2.a

Text here
```{r}
#code here
```

## Exercise 2.2.b

Text here

```{r}
#code here
```

## Exercise 2.2.c

Text here

```{r}
#code here
```


# Exercise 2.3
## Exercise 2.3.a

The decision variables for this optimization problem are defined as quantities of the following products:

$x_1 = Carrots; x_2 = Potatos; x_3 = Bread; x_4 = Cheese; x_5 = PeanutButter$

Each product has the corresponding nutritional values: $calories_i; fat_i; protein_i; carbs_i$
These nutritional values are used to define the decision variables below:

1. $(\sum_{i=1}^{n=5} calories_i*x_i ) \geq 2000$
1. $(\sum_{i=1}^{n=5} fat_i*x_i )  \geq 50$
1. $(\sum_{i=1}^{n=5} protein_i*x_i )  \geq 250$
1. $(\sum_{i=1}^{n=5} carbs_i*x_i)  \geq 200$


Objective is to minimize the prize:

$min(\sum_{i=1}^{n=5} price_i*x_i)$

The cheapest diet that fulfills these constraints consists out of $7.72$ servings of potatos and $9.28$ servings of peanut butter, and has a prize of $2.32$ per serving.


## Exercise 2.3.b

I think my calculations for this exercise in 02_assignment_02_03.xslxare not correct yet :(


## Exercise 2.3.c

Using an ILO model, we add the constraint of using no fractions when calculating the cheapest diet.
The cheapest diet consists out of $9$ servings of potatos and $9$ servings of peanut butter, and has a prize of $2.43$ per serving. More potatos are now required to compensate for the rounded down amount of peanut butter servings.

# Exercise 2.4
## Exercise 2.4.a

The transportation problem was solved using a LO Model. The cheapest transportation plan that fuflfills all demand and respects available supplies costs 460€

Constraints are defined as:

1. $x_{11} + x_{12} + x_{13} + x_{14} \leq 20$
1. $x_{21} + x_{22} + x_{23} + x_{24} \leq 25$
1. $x_{31} + x_{32} + x_{33} + x_{34} \leq 15$
1. $x_{11} + x_{21} + x_{31} \geq 10$
1. $x_{12} + x_{22} + x_{32} \geq 15$
1. $x_{13} + x_{23} + x_{33} \geq 15$
1. $x_{14} + x_{24} + x_{34} \geq 20$
1. $x_{ij} \geq 0, i = 1,2,3, j = 1,2,3,4$

Objective is to find the cheapest transportation plan, such that:

$min 10 x_{11} + 0x{12} + 20x{13} + 11x{14} \\ + 12x{21} + 7x{22} + 9x{23} + 20x{24} \\ + 0x{31} + 14x{32} + 16x{33} + 18x{34}$


## Exercise 2.4.b

Each transportation edge has an initial cost of 100. 

For this we will use the same constraints as in 2.4.a plus a decision contraint:

$$
y_{ij} = 
\begin{cases} 
  1 &\text{if } x_{ij} > 0\\
  0 &\text{else } 
\end{cases}
$$

The objective stays also the same but also incorporates the additional cost for transporting goods from a source to a destination.

$min \displaystyle\sum_{i=1}^{n=3} \displaystyle\sum_{j=1}^{m=4} c_{ij} x_{ij} + 100y_{ij}$

where $y_{ij}$ is a binary decision variable whether edge $x_{ij}$ has been used or not and adds the additional cost respectively.

With these fixed costs, the cheapest price for the transportation of all goods to the requested destinations is 1060€

# Exercise 2.5
## Exercise 2.5.a
To find the cheapest solution to this personal scheduling problem, an ILO model was used. The ILO model is de the best model to use because there can't be any half-employees. After running the excel optimization solver, the cheapest schedule would be 3,296.00 euros. The objective and constraint can be found below. The full excel sheet can be found in the Appendix. Here it can be seen when the shift needs to start and at what time.

__Outline of the ILO model:__ \

2 types of shifts: \

* 4-hour, 24 euro per hour (96 euro total)
* 8-hour (not inlcluding break of 0.5 hour), 20 euro per hour (160 euro total)

__Objective:__ (minimize the costs) \
min(∑(number of 4 hour shifts) * 96 + ∑(number of 8 hour shifts) * 160) \

__Decision variable:__ \
x~i~ = number of staffers (4 or 8 hour shift) per half hour \

__Constraints:__
The number of worker had to be Scheduled~u~ >= Required~u~ \
i = 1, 2, ..., 25 \
u = 1, 2, ..., 24 \
x~i~ ∈ 0, 1, 2, ...

## Exercise 2.5.b
In the new problem, two things changed compared to the previous problem. First, now there were only 8-hour shifts. Second, the "required staffing" is now "demanded staffing", suggesting that it is not necessarily to be fulfilled anymore. This changed the layout of the excel sheet together with several constraints. The excel sheet is displayed in the Appendix; the ILO model is presented below.

__Objective:__ (minimizes the sum of absolute differences between the demanded and scheduled staffing) \
min(∑(demanded staffing for each half hour) - ∑(scheduled staffing for each half hour))

__Decision variable:__ \
x~i~ = number of scheduled staffing per half hour. Earliest start 9:00 latest start 12:30.

__Constraints:__ \
The number of Scheduled workers had to be smaller or equal than the Demanded workers:\
workers had to be Scheduled~u~ <= Demanded~u~ \
i = 1, 2, ..., 8 \
u = 1, 2, ..., 24 \
x~i~ ∈ 0, 1, 2, ...
