---
title: "Assignment 2 - Group 4"
author: "Katrin Grunert (13914103), Carmen Byrne Salsas (12968900), Bono Lardinois (11713364)"
date: "14 October 2022"
output:
  pdf_document: default
fontsize: 11pt
highlight: tango
---

`r sat = read.table("sat.txt",header=TRUE)`
`r tree = read.table("treeVolume.txt",header=TRUE)`

# Exercise 2.1

## Exercise 2.1.a

Text here

```{r,echo=FALSE,fig.height=3.5}
#code here
```

## Exercise 2.1.b

Text here

```{r, echo=FALSE}
#code here
```

## Exercise 2.1.c

Text here

```{r}
#code here
```
## Exercise 2.1.d

Text here

```{r}
#code here
```

# Exercise 2.2
## Exercise 2.2.a

Text here
```{r}
#code here
```

## Exercise 2.2.b

Text here

```{r}
#code here
```

## Exercise 2.2.c

Text here

```{r}
#code here
```


# Exercise 2.3
## Exercise 2.3.a

The decision variables for this optimization problem are defined as quantities of the following products:

$x_1 = Carrots; x_2 = Potatos; x_3 = Bread; x_4 = Cheese; x_5 = PeanutButter$

Each product has the corresponding nutritional values: $calories_i; fat_i; protein_i; carbs_i$
These nutritional values are used to define the decision variables below:

1. $(\sum_{i=1}^{n=5} calories_i*x_i ) \geq 2000$
1. $(\sum_{i=1}^{n=5} fat_i*x_i )  \geq 50$
1. $(\sum_{i=1}^{n=5} protein_i*x_i )  \geq 250$
1. $(\sum_{i=1}^{n=5} carbs_i*x_i)  \geq 200$


Objective is to minimize the prize:

$min(\sum_{i=1}^{n=5} price_i*x_i)$

The cheapest diet that fulfills these constraints consists out of $7.72$ servings of potatos and $9.28$ servings of peanut butter, and has a prize of $2.32$ per serving.


## Exercise 2.3.b

Text here

```{r}
#code here
```


## Exercise 2.3.c

Text here

```{r}
#code here
```

# Exercise 2.4
## Exercise 2.4.a

Text here

```{r}
#code here
```

## Exercise 2.4.b

Text here

```{r}
#code here
```

# Exercise 2.5
## Exercise 2.5.a

Text here

```{r}
#code here
```

## Exercise 2.5.b

Text here

```{r}
#code here
```
