---
title: "Assignment 2 - Group 4"
author: "Katrin Grunert (13914103), Carmen Byrne Salsas (12968900), Bono Lardinois (11713364)"
date: "14 October 2022"
output:
  pdf_document: default
fontsize: 11pt
highlight: tango
---

```{r,echo=FALSE,fig.height=3.5}
sat = read.table("sat.txt",header=TRUE)
tree = read.table("treeVolume.txt",header=TRUE)
```

# Exercise 2.1

## Exercise 2.1.a

Text here

```{r,echo=FALSE,fig.height=3.5}
#code here
```

## Exercise 2.1.b

Text here

```{r, echo=FALSE}
#code here
```

## Exercise 2.1.c

Text here

```{r}
#code here
```
## Exercise 2.1.d

Text here

```{r}
#code here
```

# Exercise 2.2
## Exercise 2.2.a

Text here
```{r}
#code here
```

## Exercise 2.2.b

Text here

```{r}
#code here
```

## Exercise 2.2.c

Text here

```{r}
#code here
```


# Exercise 2.3
## Exercise 2.3.a

The decision variables for this optimization problem are defined as quantities of the following products:

$x_1 = Carrots; x_2 = Potatos; x_3 = Bread; x_4 = Cheese; x_5 = PeanutButter$

Each product has the corresponding nutritional values: $calories_i; fat_i; protein_i; carbs_i$
These nutritional values are used to define the decision variables below:

1. $(\sum_{i=1}^{n=5} calories_i*x_i ) \geq 2000$
1. $(\sum_{i=1}^{n=5} fat_i*x_i )  \geq 50$
1. $(\sum_{i=1}^{n=5} protein_i*x_i )  \geq 250$
1. $(\sum_{i=1}^{n=5} carbs_i*x_i)  \geq 200$


Objective is to minimize the prize:

$min(\sum_{i=1}^{n=5} price_i*x_i)$

The cheapest diet that fulfills these constraints consists out of $7.72$ servings of potatos and $9.28$ servings of peanut butter, and has a prize of $2.32$ per serving.


## Exercise 2.3.b

I think my calculations for this exercise in 02_assignment_02_03.xslxare not correct yet :(


## Exercise 2.3.c

Using an ILO model, we add the constraint of using no fractions when calculating the cheapest diet.
The cheapest diet consists out of $9$ servings of potatos and $9$ servings of peanut butter, and has a prize of $2.43$ per serving. More potatos are now required to compensate for the rounded down amount of peanut butter servings.

# Exercise 2.4
## Exercise 2.4.a

The transportation problem was solved using a LO Model. The cheapest transportation plan that fuflfills all demand and respects available supplies costs 460€

Constraints are defined as:

1. $x_{11} + x_{12} + x_{13} + x_{14} \leq 20$
1. $x_{21} + x_{22} + x_{23} + x_{24} \leq 25$
1. $x_{31} + x_{32} + x_{33} + x_{34} \leq 15$
1. $x_{11} + x_{21} + x_{31} \geq 10$
1. $x_{12} + x_{22} + x_{32} \geq 15$
1. $x_{13} + x_{23} + x_{33} \geq 15$
1. $x_{14} + x_{24} + x_{34} \geq 20$
1. $x_{ij} \geq 0, i = 1,2,3, j = 1,2,3,4$

Objective is to find the cheapest transportation plan, such that:

$min 10 x_{11} + 0x{12} + 20x{13} + 11x{14} \\ + 12x{21} + 7x{22} + 9x{23} + 20x{24} \\ + 0x{31} + 14x{32} + 16x{33} + 18x{34}$


## Exercise 2.4.b

Each transportation edge has an initial cost of 100. 

For this we will use the same constraints as in 2.4.a plus a decision contraint:

$$
y_{ij} = 
\begin{cases} 
  1 &\text{if } x_{ij} > 0\\
  0 &\text{else } 
\end{cases}
$$

The objective stays also the same but also incorporates the additional cost for transporting goods from a source to a destination.

$min \displaystyle\sum_{i=1}^{n=3} \displaystyle\sum_{j=1}^{m=4} c_{ij} x_{ij} + 100y_{ij}$

where $y_{ij}$ is a binary decision variable whether edge $x_{ij}$ has been used or not and adds the additional cost respectively.

With these fixed costs, the cheapest price for the transportation of all goods to the requested destinations is 1060€

# Exercise 2.5
## Exercise 2.5.a

Text here

```{r}
#code here
```

## Exercise 2.5.b

Text here

```{r}
#code here
```
