---
title: "Assignment 1"
author: "Katrin Grunert, Carmen Byrne, Bono ???"
date: "16 September 2022"
output: pdf_document
fontsize: 11pt
highlight: tango
---

# Exercise 1.1

## Exercise 1.1.a

`r birthweight = read.table("birthweight.txt",header=TRUE)`

The point estimate for $\mu$ is `r round(mean(birthweight$birthweight), 3)`.

The p-value `r round(shapiro.test(birthweight$birthweight)$p.value, 4)` of the Shapiro-Wilk-Test suggests, that the hypothesis that the data is normally distributed is not rejected. 

Histogram and QQPlot of the data also suggest, that the data is approximately normally distributed


```{r,echo=FALSE,fig.height=3.5}
par(mfrow=c(1,2))
qqnorm(birthweight$birthweight)
qqline(birthweight$birthweight, col = "red")
hist(birthweight$birthweight)
```

## Exercise 1.1.b

```{r,echo=FALSE}
n = nrow(birthweight); mu = mean(birthweight$birthweight); s = sd(birthweight$birthweight)
error_margin = qt(0.90,df=n-1)*s/sqrt(n)

low = mu - error_margin 
up = mu + error_margin
```

n = `r round(n, 3)`; $\mu$ = `r round(mu, 3)`; s = `r round(s, 3)`

The 90% Confidence interval is [`r round(low, 3)`, `r round(up, 3)`]

## Exercise 1.1.c
```{r}
t.test(birthweight$birthweight, mu = 2800, alternative = "greater")
```

## Exercise 1.1.d
The Student T-Test from Exercise 1.1.c is one-sided, because we were only interested whether the birthweight mean is bigger than 2800 or not.


# Exercise 1.2
## Exercise 1.2.a
## Exercise 1.2.b
## Exercise 1.2.c
## Exercise 1.2.d


# Exercise 1.3
## Exercise 1.3.a
## Exercise 1.3.b
## Exercise 1.3.c
## Exercise 1.3.d

# Exercise 1.4
## Exercise 1.4.a
## Exercise 1.4.b
## Exercise 1.4.c
## Exercise 1.4.d