---
title: "Assignment 1 - Group 4"
author: "Katrin Grunert, Carmen Byrne, Bono Lardinois"
date: "22 September 2022"
output: pdf_document
fontsize: 11pt
highlight: tango
---

`r weather = read.table("weather.txt",header=TRUE)`
`r birthweight = read.table("birthweight.txt",header=TRUE)`

# Exercise 1.1

## Exercise 1.1.a


The point estimate for the population mean is `r round(mean(birthweight$birthweight), 3)`.

The p-value `r round(shapiro.test(birthweight$birthweight)$p.value, 4)` of the Shapiro-Wilk-Test suggests, that the hypothesis that the data is normally distributed is not rejected. 

QQPlot and Histogram of the data also suggest, that the data is approximately normally distributed


```{r,echo=FALSE,fig.height=3.5}
par(mfrow=c(1,2))
qqnorm(birthweight$birthweight, main="QQPlot of Birthweight")
qqline(birthweight$birthweight, col = "red")
hist(birthweight$birthweight, main="Histogram of Birthweight")
```

## Exercise 1.1.b

```{r,echo=FALSE}
n = nrow(birthweight); mu = mean(birthweight$birthweight); s = sd(birthweight$birthweight)
error_margin = qt(p=(1-0.9)/2,df=n-1, lower.tail = FALSE)*s/sqrt(n)

low = mu - error_margin 
up = mu + error_margin
```

sample size n = `r round(n, 3)`; population $\mu$ = `r round(mu, 3)`; sample standard deviation = `r round(s, 3)`; $\alpha$ = 0.05 (90% CI two-sided)

The 90% Confidence interval has the lower and upper bounds of [`r round(low, 3)`, `r round(up, 3)`], meaning we can say with 90% confidence that the population mean lies between these values. 

## Exercise 1.1.c
```{r}
t.test(birthweight$birthweight, mu = 2800, alternative = "greater")
```

## Exercise 1.1.d
The Student T-Test from Exercise 1.1.c is one-sided, because we were interested whether the birthweight mean is bigger than 2800, resulting in a right-tailed and not a two-sided test.


# Exercise 1.2
## Exercise 1.2.a
## Exercise 1.2.b
## Exercise 1.2.c
## Exercise 1.2.d


# Exercise 1.3
## Exercise 1.3.a
```{r}
summary(weather$humidity)
summary(weather$temperature)
```

```{r,echo=FALSE,fig.height=3.5}
par(mfrow=c(1,2))
boxplot(weather$humidity, main='Boxplot of Humidity')
boxplot(weather$temperature, main='Boxplot of Temperature')
```
Summary and boxplot of the temperature data show a greater range of values (see Min and Max values)
In both summaries of humidity and temperature, no severe outliers can be found, as Mean and Median are close to each other and no outliers are being displayed in the boxplots.

When plotting both data in a scatterplot (below), no clear correlation can be found between humidity and temperature.

```{r,echo=FALSE,fig.height=4.5, fig.width=4.5}
par(mfrow=c(1,1))
plot(weather$humidity, weather$temperature, xlab='Humidity', ylab="Temperature", main="Scatterplot of HumidityxTemperature")
```


## Exercise 1.3.b

Histogram and QQPlot of the temperature data suggest that the data is not normally distributed, neither a clear bell curve can be drawn into the histogram, nor a straight line can be fitted to approximately all data points in the QQPlot.

```{r,echo=FALSE,fig.height=3.5}
par(mfrow=c(1,2))
hist(weather$temperature, main = 'Histogram of Temperature')
qqnorm(weather$temperature, main = 'QQPlot of Temperature')
qqline(weather$temperature, col = "red")
```


## Exercise 1.3.c

```{r,echo=FALSE}
n = nrow(weather); mu = mean(weather$temperature); s = sd(weather$temperature)
error_margin = qt((1-0.90)/2,df=n-1, lower.tail=FALSE)*s/sqrt(n)

low = mu - error_margin 
up = mu + error_margin
```

Assuming normality of the data, we can say with 90% confidence that the mean temperature of the population lays in the range of [`r round(low,3)`, `r round(up, 3)`]

## Exercise 1.3.d

# Exercise 1.4
## Exercise 1.4.a
The chi-square test of homogeneity tests whether two datasets are coming from the same population. This seems to fit to the question how well the author imitates Austen's style of writing.

## Exercise 1.4.b
## Exercise 1.4.c
## Exercise 1.4.d