---
title: 'Assignment 1 '
author: "Bono Lardinois"
date: "9/15/2022"
output:
  html_document: default
  pdf_document: default
---
```{r}
set.seed(123)
```

## Child benefit
From a fictive sample (n = 200) of working parents, 140 parents receive childcare benefits. This report section will investigate the fraction (p) of working parents that receive childcare benefits—starting with calculating the point estimate for p. The point estimate of p is the chance of success. In this case, success means having child benefits. This is done by dividing the parents with child benefits (140) by the total sample size (200). In this case, this will be:\n
p-hat = 140/200 = 0.7

```{r, echo=FALSE}
n = 200
# p 
p = 140/n
# q 
q = 1 - p

```

To derive a 99% confidence interval for the p-hat we frist need to calculate the chance of failure (q). To estimate the chance of failure (no child benefit), we use the formula q = 1 - p. In this case, the q is 0.3.\n
Next, we can derive the z-alpha with the q-norm function from R. When we fill in the alpha level of 0.99, we get a z-alpha of 2.326.
```{r, results='hide'}
z_alpha = qnorm(0.99)
round(z_alpha, 3)
```

With the z-alpha, the p-value, the q-value, and the sample size, we can calculate the upper and lower bound of the 99% confidence interval. This results in a lower-bound of 0.662 and an upper bound of 0.738. 
```{r, results='hide'}
# upperbound
upper = p + z_alpha/2*sqrt((p*q)/n)
# lowerbound
lower = p - z_alpha/2*sqrt((p*q)/n)
CI = c(lower, upper)
print(round(CI, 3))
```

Finally, we will test whether the null hypothesis that the fraction is equal to 75% holds with an α of 0.1. Meaning H0: p = 0.75, H1: p ≠ 0.75.\n
To test this, we will make use of a binomial test. The binomial test can be used to determine whether the population proportion of one level in a dichotomous variable equals a specific claimed value. In our case, we are dealing with a dichotomous variable because parents either have child benefits or do not. \n
For the binomial test, we need the following variables: number of success (140), sample size (200), and p-value: in this case, p = p0 = 0.75 and a confidence level of 1 - α. \n 
The first test was conducted with an alpha level of 0.1, conveying a confidence interval (CI) of (1 - 0.1) 90%. This test gave us a p-value of 0.103 and a confidence interval of [0.642, 0.753]. Meaning that 75% would be inside of the confidence interval. \n
The second test used an alpha of 0.15 (CI of 85%). Resulting in the same p-value and an interval of [0.649, 0.747]. \n
We used an alpha of 0.2 (CI of 80%) for the third test. Here we got the same p-value as before and a confidence interval of [0.655, 0.742]. Implying that 75% would be outside of the CI with an alpha level of 0.2.

```{r}
test_90 = binom.test(140,200, p = 0.75, conf.level = 0.9)
test_85 = binom.test(140,200, p = 0.75,conf.level = 0.85)
test_80 = binom.test(140,200, p = 0.75,conf.level = 0.8)

tab <- matrix(c(test_90$conf.int, test_85$conf.int, test_80$conf.int, test_90$conf.int[2], test_85$conf.int[2], test_80$conf.int[2]), ncol=2,nrow = 3, byrow=TRUE)
colnames(tab) <- c("lower", "Upper")
rownames(tab) <- c('Alpha 0.1','Alpha 0.15','Alpha 0.2')
tab <- as.table(tab)
tab

```

## Jane Austen
We can see four columns and six rows in the dataset described below. The respective columns mean the following: \n
* Sense is chapters 1 and 3 of the book Sense and Sensibility by Jane Austen.
* Emma is chapters 1,2, and 3 from the book Emma by Jane Austen.
* Sand1 is chapters 1 and 6 of Sanditon by Jane Austen.
* Sand2 is chapters 12 and 24, Sanditon by an admirer. 
```{r}
austen <- read.table("~/Desktop/SSO/Week_3/austen.txt")
austen
```

#### Testing for Independence or Homogeneity
Testing for independence or homogeneity depends on what you want to know. With independence testing, there is one large sample, and then you are testing whether the row and column variables differ significantly. With testing for homogeneity, samples are taken from different populations, with one sample per column. Here we test whether the columns differ significantly from each other. In our case, we want to test for consistency between the different samples, so we want to see whether the columns are the same or differ. We will use the Homogeneity test to look for a difference between the columns.\n
\n
To test whether Austen was consistent in her use of words throughout the writing of her books, we performed a chi-squared test. The H0, in this case, is that row categories are the same for each column. The H1 is that the distribution among the row categories is not the same for each column.\n
The chi-squared test is used to compare observed results and expected results. It has its purpose of discovering whether there is a significant difference between the observed and the expected data.\n
To test only the books/chapter of Austen, we created a subset with only the books/chapters that were written by Austen. With this new dataset, we performed a chi-squared test. This gives us a p-value of 0.161. This p-value is > 0.05, meaning we cannot reject the null hypotheses and state that there is a significant difference between the different columns (books/chapters) in the data frame. 
```{r}
df_1 <- austen[c("Sense","Emma", "Sand1")]
austen_chi_1 = chisq.test(df_1)
round(austen_chi_1$p.value, 3)
```
To check the condition of the chi-squared test: At least 80% of the expected values should be at least 5. We took the expected table from the chi-squared test and made a boolean statement (Expected > 5). This table showed that every cell has a value greater than 5, meaning the condition is met.  
```{r}
austen_chi_1$expected > 5
```
To check whether the admirer was any good at imitating Austen's writing style, we can compare the chapters from the Sandition book. For this, we made a new data frame and performed another chi-squared. Here the p-value was 0.131. This p-value is greater than 0.05, meaning we can not say there is a significant difference between the chapters written by Austen and her admirer based on the given data. Moreover the condition of the chi-squared was met because all cells of the expected table were greater than 5. 
```{r}
df_2 <- austen[c("Sand1", "Sand2")]
austen_chi_2 = chisq.test(df_2)
round(austen_chi_2$p.value, 3)
```
```{r}
austen_chi_2$expected > 5
```









